### 2024-03-01 manuscript.tex
Added citation for evolutionary analysis by sequence shuffling
Added commas after e.g. in some sentences

To illustrate the utility of this method for generating hypotheses to guide the functional dissection of specific proteins, we next compared two regions with elevated values of fraction\_Q and repeat\_QN but different signals of conservation for each (Fig~\ref{fig:alignment}A-B). Though the alignments of both regions are enriched in glutamine and asparagine residues (Fig~\ref{fig:alignment}C,E), the first has significant signals of conservation for both features (Fig~\ref{fig:alignment}D), whereas the second does not (Fig~\ref{fig:alignment}F). Notably, the first region is from an orthologous group that contains a sequence from the \textit{D. melanogaster} gene {ewg}, which is a transcription factor. As polyglutamine stretches can modulate the activity of transcription factor activation domains~\cite{Mitchell1989, Gerber1994}, this observation is consistent with the maintenance of an optimal level of transcriptional activation mediated by constraints on glutamine-associated features. In contrast, the low log likelihood ratios of the second region, whose \textit{D. melanogaster} gene is unannotated, are suggestive of unconstrained glutamine repeat expansion and contraction. Interestingly, this region has highly significant log likelihood ratios for several features associated with glycine or phenylalanine repeats despite having few of these residues. However, careful inspection of the alignment reveals several positions where such repeats appear in only one or two sequences, \textit{e.g.} columns 153 and 208. While these strong signals of conservation may represent artifacts of applying a continuous model of trait evolution to features derived from discrete data, they could also indicate true constraints against the presence of flexible or hydrophobic patches in those sequence segments. Taken together, these results highlight this method's ability to identify biochemical features with signals of constrained evolution while also emphasizing the importance of combining this information with a protein's biological context to jointly suggest testable hypotheses of function.

@@

\subsection{Disorder is correlated with many molecular features}
Though the mutations generated by evolution are not a random or exhaustive sample of sequence space, they are perturbations of a common ancestor which can reveal the relationship between a region's biophysical properties and its propensity for disorder. Thus, the analysis of feature and score contrasts is effectively a natural ``mutational scanning'' experiment. We found that disorder scores have the strongest correlations with features that measure a region's overall polarity and hydrophobicity, \textit{e.g.} fraction\_disorder and hydropathy. As the formation of a hydrophobic core is a major driving force in protein folding, this relationship is expected. However, the strength of these correlations indicate that a region's relative proportion of hydrophilic and hydrophobic residues is, to a first approximation, the largest determinant of predicted intrinsic disorder.

@@

While the BM and OU models are powerful tools for studying trait evolution, their generality limits the specificity of the hypotheses they can test. In contrast, because simulation-based approaches can specify arbitrary constraints, they permit investigations of increasingly refined models of IDR evolution. We therefore view the two approaches as complementary and propose the use of hybrid methods where test statistics are derived from phylogenetic comparative methods like the BM model, and simulations generate the null distributions for those test statistics. While simulations can eliminate specific hypotheses, the substantial resources involved both in designing them and generating samples make exhaustively testing mechanisms of feature constraint by simulation impractical. We instead recommend a workflow that begins with an analysis using general models of trait evolution to suggest specific hypotheses of constraint that are then tested with a simulation-based approach. Another hybrid method involves using sequence permutations to test for the conservation of motifs or patterned features like kappa. In this method, a comparative model like BM provides the test statistic as before, and a sample of shuffled sequences approximates the null distribution. Because the background distribution of residues is preserved, this procedure can specifically test for the conservation of motifs or patterned features in an IDR independent of the conservation of its composition.

>>>>

To illustrate the utility of this method for generating hypotheses to guide the functional dissection of specific proteins, we next compared two regions with elevated values of fraction\_Q and repeat\_QN but different signals of conservation for each (Fig~\ref{fig:alignment}A-B). Though the alignments of both regions are enriched in glutamine and asparagine residues (Fig~\ref{fig:alignment}C,E), the first has significant signals of conservation for both features (Fig~\ref{fig:alignment}D), whereas the second does not (Fig~\ref{fig:alignment}F). Notably, the first region is from an orthologous group that contains a sequence from the \textit{D. melanogaster} gene {ewg}, which is a transcription factor. As polyglutamine stretches can modulate the activity of transcription factor activation domains~\cite{Mitchell1989, Gerber1994}, this observation is consistent with the maintenance of an optimal level of transcriptional activation mediated by constraints on glutamine-associated features. In contrast, the low log likelihood ratios of the second region, whose \textit{D. melanogaster} gene is unannotated, are suggestive of unconstrained glutamine repeat expansion and contraction. Interestingly, this region has highly significant log likelihood ratios for several features associated with glycine or phenylalanine repeats despite having few of these residues. However, careful inspection of the alignment reveals several positions where such repeats appear in only one or two sequences, \textit{e.g.}, columns 153 and 208. While these strong signals of conservation may represent artifacts of applying a continuous model of trait evolution to features derived from discrete data, they could also indicate true constraints against the presence of flexible or hydrophobic patches in those sequence segments. Taken together, these results highlight this method's ability to identify biochemical features with signals of constrained evolution while also emphasizing the importance of combining this information with a protein's biological context to jointly suggest testable hypotheses of function.

@@

\subsection{Disorder is correlated with many molecular features}
Though the mutations generated by evolution are not a random or exhaustive sample of sequence space, they are perturbations of a common ancestor which can reveal the relationship between a region's biophysical properties and its propensity for disorder. Thus, the analysis of feature and score contrasts is effectively a natural ``mutational scanning'' experiment. We found that disorder scores have the strongest correlations with features that measure a region's overall polarity and hydrophobicity, \textit{e.g.}, fraction\_disorder and hydropathy. As the formation of a hydrophobic core is a major driving force in protein folding, this relationship is expected. However, the strength of these correlations indicate that a region's relative proportion of hydrophilic and hydrophobic residues is, to a first approximation, the largest determinant of predicted intrinsic disorder.

@@

While the BM and OU models are powerful tools for studying trait evolution, their generality limits the specificity of the hypotheses they can test. In contrast, because simulation-based approaches can specify arbitrary constraints, they permit investigations of increasingly refined models of IDR evolution. We therefore view the two approaches as complementary and propose the use of hybrid methods where test statistics are derived from phylogenetic comparative methods like the BM model, and simulations generate the null distributions for those test statistics. While simulations can eliminate specific hypotheses, the substantial resources involved both in designing them and generating samples make exhaustively testing mechanisms of feature constraint by simulation impractical. We instead recommend a workflow that begins with an analysis using general models of trait evolution to suggest specific hypotheses of constraint that are then tested with a simulation-based approach. Another hybrid method involves using sequence permutations to test for the conservation of motifs or patterned features like kappa~\cite{Cohan2022}. In this method, a comparative model like BM provides the test statistic as before, and a sample of shuffled sequences approximates the null distribution. Because the background distribution of residues is preserved, this procedure can specifically test for the conservation of motifs or patterned features in an IDR independent of the conservation of its composition.
